Home

### 2020-05-20 analysis/VEB-boost-fitting.Rmd
test mr.ash

## Problem description

$$Y_{N \times 1}= X_{N \times J}\beta_{J\times1} + G_{N\times M}\theta_{M\times1} + \epsilon_{N\times1}$$
$$ \epsilon \sim MVN(0,I_N\tau^{-1})$$
Spike and slab prior for $\beta_j$:
$$\beta_j|\pi,\tau \sim (1- \pi_\beta)\delta_0 + \pi_\beta N(0, \sigma_\beta^2/\tau)$$
Spike and slab prior for $\theta_m$:
$$\theta_m | \tau \sim (1- \pi_\theta)\delta_0 + \pi_\theta N(0, \sigma_\theta^2/\tau)$$
We want to estimate $\omega=\{\tau,\pi_\beta, \pi_\theta, \sigma_\beta^2,\sigma_\theta^2\}$

>>>>

## Basic idea of veb_boost

You want to add up 2 linear regressions, which we might call “learners”.
A learner is, roughly, a fucntion, that learns a relationshipo between Y and some predictors X (here the Xs are different for your 2 different learners) Let’s write this $Y=L1(X1) + L2(X2) + E$.

The idea is that you can fit this model iteratively, by first applying $L1$, and then applying $L2$, each time applying it to the appropriate residuals.So you apply $L1$ to learn a relationship between $R=Y-L2$ (residuals) and $X1$, and then $L2$ to learn a relationship between $R=Y-L1$ and $X2$ and you just iterate….

## Our model

Model:

$$Y = \mu_1 + \mu_2 + \epsilon$$
$$ \mu_1= X_{N \times J}\beta_{J\times1}$$
$$ \mu_2 = G_{N\times M}\theta_{M\times1}$$
$$ \epsilon \sim MVN(0,\sigma^2I_N) $$
Priors:

ASH prior for scaled $\beta_j$ and $\theta_m$
$$\beta_j | g_1, \sigma \sim g^1_\sigma(\cdot)$$
ASH prior for $\theta_m$:
$$\theta_m | g_2, \sigma \sim g^2_\sigma(\cdot)$$

## Input required by veb_boost

take $\mu1$ as the example, $\mu2$ works similarly.

* first moment of posterior variational approximation ($E_q(\mu_i)\in R^N$)

$$E_q(\mu_1) = X\bar{\beta}$$
$\bar{\beta}$ is posterior mean of $\beta$ given by `mr.ash` function.

* second moment of posterior variational approximation ($E_q(\mu^2_i)\in R^N_+$)

$$E_q(\mu_{1}^2) = \sum_{j=1}^J{x^2_j E_q(\beta^2)} + [E_q(\mu_1)]^2 =\sum_{j=1}^J\sum_{k=1}^K{x^2_j\phi_k \frac{\sigma^2 d_j^2 \sigma_k^2}{1+ d_j^2 \sigma_k^2}} + [E_q(\mu_1)]^2$$
where $$d_j=x^T_jx_j= \left\lVert x_j\right\rVert^2$$.

$d_j, \phi, \sigma, \sigma_k$ are estimates given by `mr.ash` function.

* KL-divergence between variational posterior (q) and prior (p).

Shall we just use ELBO from mr.ash?

### 2020-05-20 analysis/index.Rmd
test mr.ash

# Model
  * [Polygenic model](polygenic_model.html)
  * [Sparse model](sparse_model.html)

@@

## For a single gene
  * [Single gene](simulation-singleG.html)

## Under polygenic model
  * [Simulation: WTCCC](simulation-WTCCC.html)
  * [Simulation: WTCCC 2](simulation-WTCCC2.html)

# References
  * VEB-boost
    * [Problem description](VEB-boost-fitting.html)
    * Andrew's [software](https://github.com/stephenslab/VEB.Boost) and [workflowr website](https://andrewg3311.github.io/VEB-Boost-workflowR/VEB_Boost_Derivation.html)

>>>>

# Model
  * [Polygenic model](polygenic_model.html)
  * [Sparse model](sparse_model.html)
  * [mr.ash model](mrash_model.html)

# Data

## Expression data
  * [Fusion/TWAS gene expression models](fusion_expr.html)

@@

## Proof of concept
  * [Proof of concept for this project using a single gene's data](simulation-singleG.html)

## Test polygenic model
Using WTCCC data

  * [Test polygenic model 1](simulation-WTCCC.html)

  * [Test polygenic model 2: use imputed expression](simulation-WTCCC2.html)

## Test mr.ash
Using UKbiobank data (chr22)

  * [Test mr.ash performance: only SNP effect](simulation-mr.ash-ukbchr22.html)

## Test veb_boost
A simplified version: interative update of two mr.ash component

  * [Using some simple data](simulation_simpleveb-boost.html)

  * [Using UKbiobank data (chr22)](simulation_simpleveb-boost_ukbchr22.html)

# Related software
  * VEB-boost
    * [Problem description](VEB-boost-fitting.html)
    * Andrew's [software](https://github.com/stephenslab/VEB.Boost) and [workflowr website](https://andrewg3311.github.io/VEB-Boost-workflowR/VEB_Boost_Derivation.html)
